---
title: "Cross Validation"
output: github_document
---

```{r set up, include = FALSE}  
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = '90%'
)

theme_set(theme_classic() + theme(legend.position = 'bottom'))

options(
  ggplot2.continous.colour = 'viridis_d',
  ggplot2.continous.fill = 'viridis_d'
)

scalr_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Simulate data

```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )
```

```{r}
nonlin_df %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point()
```

## Cross validation -- by hand

Get training and testing datasets

```{r}
train_df = sample_n(nonlin_df, size = 80)
test_df = anti_join(nonlin_df, train_df, by = 'id')
```

Fit three models.

```{r}
linear_model = lm(y ~ x, data = train_df)
smooth_model = gam(y ~ s(x), data = train_df)
wiggly_model = gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

Can I see what i just did ...

```{r}
train_df %>% 
  gather_predictions(linear_model, smooth_model, wiggly_model) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = pred), color = 'red') +
  facet_grid(. ~ model)
```

Let's look at the prediction accuracy.

```{r}
rmse(linear_model, test_df)
rmse(smooth_model, test_df)
rmse(wiggly_model, test_df)
```

## Cross validation by using `modelr`

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100)
```

What is happening here...

```{r}
cv_df %>% pull(train) %>% .[[1]] %>% as_tibble()
cv_df %>% pull(test) %>% .[[1]] %>% as_tibble()
```

```{r}
cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as.tibble),
    test = map(test, as.tibble)
  )
```

Let's try to fit models and get RMSEs for them.

```{r}
cv_df = 
  cv_df %>% 
  mutate(
    linear_model = map(.x = train, ~lm(y ~ x, data = .x)),
    smooth_model = map(.x = train, ~gam(y ~ s(x), data = .x)),
    wiggly_model = map(.x = train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(.x = linear_model, .y = test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(.x = smooth_model, .y = test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(.x = wiggly_model, .y = test, ~rmse(model = .x, data = .y)),
  )
```

What do these results say about the model choices?

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

Computing averages...

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  group_by(model) %>% 
  summarize(ave_rmse = mean(rmse))
```

## Try on a real dataset.

Import the data.

```{r}
child_growth_df = 
  read_csv("./data/nepalese_children.csv") %>% 
  mutate(
    weight_cp = (weight > 7) * (weight - 7)
  )
```

weight vs arm circumference

```{r}
child_growth_df %>% 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .3)
```

Fit the models I care about.

```{r}
linear_model = lm(armc ~ weight, data = child_growth_df)
pwlin_model = lm(armc ~ weight + weight_cp, data = child_growth_df)
smooth_model = gam(armc ~ s(weight), data = child_growth_df)
```

```{r}
child_growth_df %>% 
  gather_predictions(linear_model, pwlin_model, smooth_model) %>% 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .3) +
  geom_line(aes(y = pred), color = 'red') +
  facet_grid(. ~ model)
```

Try to understand model fit using CV.

```{r}
cv_df = 
  crossv_mc(child_growth_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

See if i can fit the models to the splits

```{r}
cv_df =
  cv_df %>% 
  mutate(
    linear_model = map(.x = train, ~lm(armc ~ weight, data = .x)),
    pwlin_model = map(.x = train, ~lm(armc ~ weight, data = .x)),
    smooth_model = map(.x = train, ~gam(armc ~ s(weight), data = .x))
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(.x = linear_model, .y = test, ~rmse(model = .x, data = .y)),
    rmse_pwlin = map2_dbl(.x = pwlin_model, .y = test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(.x = smooth_model, .y = test, ~rmse(model = .x, data = .y))
  )
```

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

